Towards Optimized Keyword Hash Maps for Clojure

### 2022-08-05 paper/paper.tex
no speculation
spelling

\title{Optimized Keyword Persistent Hash Maps for Clojure}

@@

\begin{enumerate}
  \item We give an approachable walkthrough of mechanics behinds
      HAMTs.
  \item We describe the internals of Clojure's persistent HAMT implementation,
    via our pure-Clojure reimplementation.
  \item We present a port of Clojure's HAMT from Java to Clojure
    for pedagogical purposes,
    featuring trie visualization (Figure~\label{port-visualize}).
  \item We evaluate a Clojure-specific optimization for Clojure's HAMT.
  \item We speculate on future directions for HAMT's.
\end{enumerate}

@@

The source for the mask function (below),
selects partitions
by
shifting the hash's bits right until the first
5 bits is the parition needed, then the 
first 5 bits of the result
interpretted as an integer
(Appendix~\ref{jvm-bit-remark} explains \texttt{>>>}).

@@

Figure~\ref{benchmark-code} contains the benchmark
code. The purpose of the benchmark is to simulate
a set of map associations followed by many lookups,
since our optimization only triggers after an assocation
operation.
In this version of the benchmark, we perform 20
associations, each with 400,000 lookups
on keyword entries.
Each iteration has a set number of keyword entries,
and we vary the

@@

After 1,500 keys, there is an unknown problem with our optimized
implementation. Our hand-written benchmarking scheme
does not compensate for garbage collection and is rather crude,
so we would like to revisit this benchmark in the future to 
investiate this issue.

@@

Our approach to speeding up the hash-map implementation
is related to \textit{storage strategies}
by Bolz et. al~\cite{Bolz13storagestrategies}.
They observe that dynamically typed languages often use
heterogeous collections as homogeneous, and present
optimizations to take advantage of this fact.

>>>>

\title{Towards Optimized Keyword Hash Maps for Clojure}

@@

\begin{enumerate}
  \item We give an approachable walkthrough of mechanics behinds
      HAMTs.
  \item We describe the internals of Clojure's persistent HAMT implementation,
    via our pure-Clojure reimplementation.
  \item We present a port of Clojure's HAMT from Java to Clojure
    for pedagogical purposes,
    featuring trie visualization (Figure~\label{port-visualize}).
  \item We evaluate a Clojure-specific optimization for Clojure's HAMT.
  %\item We speculate on future directions for HAMT's.
\end{enumerate}

@@

The source for the mask function (below),
selects partitions
by
shifting the hash's bits right until the first
5 bits is the partition needed, then the 
first 5 bits of the result
interpreted as an integer
(Appendix~\ref{jvm-bit-remark} explains \texttt{>>>}).

@@

Figure~\ref{benchmark-code} contains the benchmark
code. The purpose of the benchmark is to simulate
a set of map associations followed by many lookups,
since our optimization only triggers after an association
operation.
In this version of the benchmark, we perform 20
associations, each with 400,000 lookups
on keyword entries.
Each iteration has a set number of keyword entries,
and we vary the

@@

After 1,500 keys, there is an unknown problem with our optimized
implementation. Our hand-written benchmarking scheme
does not compensate for garbage collection and is rather crude,
so we would like to revisit this benchmark in the future to 
investigate this issue.

@@

Our approach to speeding up the hash-map implementation
is related to \textit{storage strategies}
by Bolz et. al~\cite{Bolz13storagestrategies}.
They observe that dynamically typed languages often use
heterogenous collections as homogeneous, and present
optimizations to take advantage of this fact.
